# IE - Master in Business Analytics and Big Data
Recommendation Systems

Building a Collaborative Filtering Engine

Author: FirstName LastName

========================================================

# Initialization

Import libraries

```{r}
library(ggplot2)
```

Load the critics dataset.

```{r}
data <- read.csv('critics.csv')
```

# 1) Pearson correlation coefficient 
Calculate the Pearson correlation coefficient between Victoria and Nuria

**Expected correlation between Victoria and Nuria**: 0.3651484

```{r fig.width=7, fig.height=6}
films <- colnames(data)[-1]
users <- data$User
trans_data <- t(data[,-1])
trans_data <- as.data.frame(trans_data)
colnames(trans_data) <- users
cor(trans_data$Victoria, trans_data$Nuria, use = 'complete.obs', method = 'pearson')
```

# 2) Compare two movie critics
Using the calculated Pearson correlation in the previous section. Compare and show the movie ratings for Victoria and Nuria on a scatter plot. How is this compared with the same ratings of Maria and Nerea on a scatter plot? What about Chris and Jim? What it means?

```{r}
plot_data <- as.data.frame(cbind(films, trans_data))
rownames(plot_data) <- NULL

plot_data1 <- plot_data[, which(names(plot_data) %in% c("Victoria", "Nuria", "films")) ]
plot_data1 <- plot_data1[complete.cases(plot_data1),]
plot1 <- ggplot(plot_data1) + geom_point(aes(x = films, y = Victoria), colour=alpha('darkorchid4', 1)) +
  geom_point(aes(x = films, y = Nuria), colour=alpha('deepskyblue', 1)) + 
  labs(y = "Rating", title = "Victoria vs. Nuria") +
  theme(axis.text.x = element_text(angle = 45))
plot1

plot_data2 <- plot_data[, which(names(plot_data) %in% c("Maria", "Nerea", "films")) ]
plot_data2 <- plot_data2[complete.cases(plot_data2),]
plot2 <- ggplot(plot_data2) + geom_point(aes(x = films, y = Maria), colour=alpha('darkorchid4', 1)) +
  geom_point(aes(x = films, y = Nerea), colour=alpha('deepskyblue', 1)) + 
  labs(y = "Rating", title = "Maria vs. Nerea") +
  theme(axis.text.x = element_text(angle = 45))
plot2

plot_data3 <- plot_data[, which(names(plot_data) %in% c("Chris", "Jim", "films")) ]
plot_data3 <- plot_data1[complete.cases(plot_data3),]
plot3 <- ggplot(plot_data3) + geom_point(aes(x = films, y = Victoria), colour=alpha('darkorchid4', 1)) +
  geom_point(aes(x = films, y = Nuria), colour=alpha('deepskyblue', 1)) + 
  labs(y = "Rating", title = "Chris vs. Jim") +
  theme(axis.text.x = element_text(angle = 45))
plot3

par(mfrow = c(3,1))
plot1
plot2
plot3
```

# 3) Top Recommendations 
Get the top 5 movies recommended for Victoria. by using a weighted average of every other critic's rankings.

**Expected result**:
3.7917013044215, 'The Matrix'
3.50776533175371, 'Forrest Gump'     
3.33118834864677, 'The Sixth Sense' 
3.11491825315719, 'Shakespeare in Love'
2.9124513228665, 'Blade Runner'

```{ fig.width=7, fig.height=6}
meanVic <- mean(trans_data$Victoria, na.rm = T)
u <- 0
meanUser <- 0
for (u in 1:length(trans_data)) {
  meanUser[u] <- mean(trans_data[,u], na.rm = T)
}
meanUser_data <- as.data.frame(cbind(meanUser, colnames(trans_data)))
colnames(meanUser_data) <- c("MeanRating", "User")

f <- 0
us <- 0
rating <- 0
for (f in 1:length(trans_data)) {
  for (us in 1:length(trans_data)) {
  rating[f] <- meanVic + (sum((trans_data[f,us] - meanUser_data[u,1]) * simil[us,1]) / simil[us,1])
  }
}
```

# 4) Top similar critics
Return the top 5 critics that are most similar to Victoria. 

**Expected result**:
0.9449112, 'Rachel'
0.5976143, 'Ana' 
0.5789794, 'Oriol'
0.4925922, 'Maria'
0.4273247, 'Carles'

```{r fig.width=7, fig.height=6}
a <- 0
correlation <- 0
for (a in 1:length(trans_data)) {
  correlation[a] <- cor(trans_data$Victoria, trans_data[,a], use = 'complete.obs', method = 'pearson')
}
simil <- as.data.frame(cbind(correlation,colnames(trans_data)))
colnames(simil) = c("Correlation", "User")
simil <- simil[-15,]
simil2 <- simil[order(simil$Correlation, decreasing = T),]
simil2[1:5,]
```